Reproducible Research: Peer Assessment 1
=========================================

## Loading and preprocessing the data  
Please make sure you set the "activity.zip" is in the working directory.
```{r}
library(ggplot2)
unzip("activity.zip")
data<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?  
1. Make a histogram of the total number of steps taken each day.
```{r}
##sum the number of steps per day
steps<-tapply(data$steps,data$date,sum)
##output histograms
hist(steps,main="Histogram of Number of Steps Taken Each Day",xlab="Steps")

```
2. Calculate and Report the mean and median total number os steps taken per day.
```{r}
mean_steps<-mean(steps,na.rm=T)
median_steps<-median(steps,na.rm=T)

mean_steps
median_steps
```
The mean number of steps is `r mean_steps` and the median is `r median_steps`.

## What is the average daily activity pattern?  
1. Make a time series plot of the 5-minute interval (x-axis) and the average 
number of steps taken, averaged across all days(x-axis).
```{r}
## average the steps accross all days
interval_steps<-tapply(data$steps,data$interval,mean,na.rm=T)
##modify the variable to make is suitable for plotting
data_2<-data.frame(interval_steps)
data_2$intervals<-rownames(data_2)
row.names(data_2)=NULL
plot(data_2$intervals,data_2$interval_steps,type="l",
     main="Average Steps vs. Interval",
      xlab="Interval Identifier",
      ylab="Average Steps")
```

2. Which 5-minute interval, on average accross all days in the dataset, 
contains the maximum number of steps?

```{r}
max_steps<-max(data_2$interval_steps)
max_interval<-data_2[data_2$interval_steps==max_steps,2]
max_steps
as.numeric(max_interval)
```
 
Thus, the interval with the highest average steps is `r max_interval` with 
`r max_steps` number of steps.

## Imputing missing values  

1. Calculate and report the total number of missing values in the dataset.

```{r}
##returns True for "NA's" and False for complete observations
na_vector<-is.na(data$steps)
number_na<-sum(na_vector)
number_na
```
The number of missin values in this data set is `r number_na`.

2. Devise a strategy for filling in all the missing values in the data. 
This strategy does not need to be sophisticated. I used means for the five
minute intervals to replace the missing values. I matched the five minute interval
and replaced the corresponind steps in the new data set with the average mean
for accross all days for that 5-minute interval.

3. Create a new dataset that is equal to the original dataset 
but with the missing data filled in. I also prinited the first ten values of the 
data set to show that the NA's are gone. 
```{r}
data_3<-data
for (i in 1:length(na_vector)) {
    if (na_vector[i]==T) {
        data_3$steps[i]<-interval_steps[[as.character(data_3$interval[i])]]
        data_3$steps[i]<-round(data_3$steps[i],0)
    }
}

data_3[1:10,]
```

4. Make a histogram of the total number of steps taken each day and Calculate and 
report the mean and median total number of steps taken per day.

```{r Histogram}
##sum the number of steps per day
steps<-tapply(data_3$steps,data_3$date,sum)
##output histograms
hist(steps,main="Histogram of Number of Steps Taken Each Day",xlab="Steps")

```

```{r Mean/Median}
mean_steps_2<-mean(steps)
median_steps_2<-median(steps)

mean_steps_2
median_steps_2
```

Yes the new values do have some impact on the estimates of the total daily number
of steps.   
```{r Comparison}
comp_table<-data.frame(c(mean_steps,median_steps),c(mean_steps_2,median_steps_2))
rownames(comp_table)<-c("mean","median")
colnames(comp_table)<-c("old","new")
comp_table
```

* the new mean is slightly lower then bofore: old -`r mean_steps` vs new 
`r mean_steps_2`.   
* the median is also lower: old `r median_steps` vs new `r median_steps_2`
* getting rid of the missing values did not change the skewness, as the 
median is lower than the the mean in both cases.That is in both cases
the distribution is negatively skewed. 

## Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels -- "weekday" 
and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
data_3$date<-weekdays(as.Date(data_3$date))
day_factor<-as.numeric((data_3$date=="Saturday")|(data_3$date=="Sunday"))
day_factor<-factor(day_factor,labels=c("Weekday","Weekend"))
data_3$date<-day_factor

data_4<-aggregate(steps~date+interval,mean,data=data_3)
```

2.Make a panel plot containing a time series plot (i.e. type = "l") 
of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis). 

```{r}

library(ggplot2)
ggplot(data=data_4, aes(x=interval, y=steps, group=1)) + geom_line()+facet_grid(date~.)





```

