Reproducible Research: Peer Assessment 1
=========================================

## Loading and preprocessing the data  
Please make sure you set the "activity.zip" is in the working directory.
```{r}
unzip("activity.zip")
data<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?  
1. Make a histogram of the total number of steps taken each day.
```{r}
##sum the number of steps per day
steps<-tapply(data$steps,data$date,sum)
##output histograms
hist(steps,main="Histogram of Number of Steps Taken Each Day",xlab="Steps")

```
2. Calculate and Report the mean and median total number os steps taken per day.
```{r}
mean_steps<-mean(steps,na.rm=T)
median_steps<-median(steps,na.rm=T)

mean_steps
median_steps
```
The mean number of steps is `r mean_steps` and the median is `r median_steps`.

## What is the average daily activity pattern?  
1. Make a time series plot of the 5-minute interval (x-axis) and the average 
number of steps taken, averaged across all days(x-axis).
```{r}
## average the steps accross all days
interval_steps<-tapply(data$steps,data$interval,mean,na.rm=T)
##modify the variable to make is suitable for plotting
data_2<-data.frame(interval_steps)
data_2$intervals<-rownames(data_2)
row.names(data_2)=NULL
plot(data_2$intervals,data_2$interval_steps,type="l",
     main="Average Steps vs. Interval",
      xlab="Interval Identifier",
      ylab="Average Steps")
```

2. Which 5-minute interval, on average accross all days in the dataset, 
contains the maximum number of steps?

```{r}
max_steps<-max(data_2$interval_steps)
max_interval<-data_2[data_2$interval_steps==max_steps,2]
max_steps
as.numeric(max_interval)
```
 
Thus, the interval with the highest average steps is `r max_interval` with 
`r max_steps` number of steps.

## Imputing missing values  

1. Calculate and report the total number of missing values in the dataset.

```{r}
##returns True for "NA's" and False for complete observations
na_vector<-is.na(data$steps)
number_na<-sum(na_vector)
number_na
```
The number of missin values in this data set is `r number_na`.

2. Devise a strategy for filling in all the missing values in the data. 
This strategy does not need to be sophisticated. I used means for the five
minute intervals to replace the missing values.

3. Create a new dataset that is equal to the original dataset 
but with the missing data filled in.
```{r}
data_3<-data
for (i in 1:length(na_vector)) {
    if (na_vector[i]==T) {
        data_3$steps[i]<-interval_steps[[as.character(data_3$interval[i])]]
        
    }
    
}
```


## Are there differences in activity patterns between weekdays and weekends?